---
title: "WQD7004_project_part"
author: "Sakshi"
date: "6/14/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### BMR Calculation

```{r}
cal_bmr = function(gen,wt,ht,age){
  if(gen == 'M'){
    66 + (13.7*wt)+5*(ht)-(6.8*age)
  }else if(gen == "F"){
    655 +(9.6*wt)+1.8*(ht)-(4.7*age)
  }else{
    print("wrong input")
  }
}
# Input required
weight = 65 
height = 152 
age = 27 
gender = 'F' 

BMR = cal_bmr(gender,weight,height,age)
BMR

```



## Activity Multiplier
```{r}

activity_cal = function(activity){
  if(activity == 'sendetary'){
  multiplier = BMR*1.2
  }else if(activity == 'lightly_active'){
  multiplier = BMR*1.375
  }else if(activity == 'moderatly_active'){
  multiplier = BMR * 1.55
  }else if(activity == 'very_active'){
  multiplier = BMR *1.725
  }else if(activity == 'extra_active'){
  multiplier = BMR*1.9
  }else{
    print("enter correct input")
  }
}

calorie_required_based_on_activity = activity_cal('lightly_active')  ## just an example
calorie_required_based_on_activity

```


## Calorie need as per goal
```{r}

goal_cal = function(goal){
  
  if(goal == "maintain_weight"){
    
    maintain_weight = calorie_required_based_on_activity
    
    carbs_cal = (50/100)*maintain_weight
    protein_cal = (30/100)*maintain_weight
    fat_cal = (20/100)*maintain_weight
    
    return(data.frame(carbs_cal,protein_cal,fat_cal))
    
  }else if(goal == "loose_weight"){
    
    loose_weight = calorie_required_based_on_activity*0.9
    
    carbs_cal = (40/100)*loose_weight
    protein_cal = (30/100)*loose_weight
    fat_cal = (30/100)*loose_weight
    
    return(data.frame(carbs_cal,protein_cal,fat_cal))
    
  }else if(goal == "gain_weight"){
    
    gain_weight = calorie_required_based_on_activity *1.1
    
    carbs_cal = (55/100)*gain_weight
    protein_cal = (30/100)*gain_weight
    fat_cal = (15/100)*gain_weight
    
    return(data.frame(carbs_cal,protein_cal,fat_cal))
    
  }else if(goal=="RDA Typical"){
    
     maintain_weight = calorie_required_based_on_activity
     
     carbs_cal = (50/100)*maintain_weight
     protein_cal = (25/100)*maintain_weight
     fat_cal = (25/100)*maintain_weight
     return(data.frame(carbs_cal,protein_cal,fat_cal))
     
  }else if(goal=="Muscle gain"){
    
    maintain_weight = calorie_required_based_on_activity
    
    carbs_cal = (55/100)*maintain_weight
    protein_cal = (30/100)*maintain_weight
    fat_cal = (30/100)*maintain_weight
    return(data.frame(carbs_cal,protein_cal,fat_cal))
    
  }else if(goal=="keto diet"){
    
    maintain_weight = calorie_required_based_on_activity
    
    carbs_cal = (8/100)*maintain_weight
    protein_cal = (20/100)*maintain_weight
    fat_cal = (65/100)*maintain_weight
    return(data.frame(carbs_cal,protein_cal,fat_cal))
  }else{
    
    print("wrong input")
  }
  
}
micronutrients = goal_cal("Muscle gain")
micronutrients

```

## Loading the data set and modifying grams into calories
```{r}
my_nurients_data = read.csv("C:/Users/SAKSHI/Documents/R/WQD7001_Project/nutrients_cleaned.csv",stringsAsFactors = FALSE)

## error in the values of oyster
my_nurients_data = my_nurients_data %>%  mutate(Calories = replace(Calories,Calories==231,458),
                                               Carbs = replace(Carbs,Carbs==236,28),
                                               Protein = replace(Protein,Protein == 232,21),
                                               Fat = replace(Fat,Fat==233,23),
                                               Sat.Fat = replace(Sat.Fat,Sat.Fat==234,8),
                                               Fiber=replace(Fiber,Fiber==235.0,0.0))


## Calculating Nutrients per gram

```{r}
nutrients_per_gram = my_nurients_data %>% mutate(carbs_per_gram = Carbs/Grams,
                                                 protein_per_gram = Protein/Grams,
                                                 fat_per_gram = Fat/Grams) %>% 
  select(Food,Grams,Carbs,Protein,Fat,carbs_per_gram,protein_per_gram,fat_per_gram)
#nutrients_per_gram

max_nutrients = nutrients_per_gram

max_carbs = max(max_nutrients$carbs_per_gram)
max_carbs
highest_carbs= max_nutrients %>% filter(carbs_per_gram == max_carbs) %>% select(Food,Grams,Carbs,Protein,Fat,carbs_per_gram,protein_per_gram,fat_per_gram)

highest_carbs # food with max carbs

#-------------------------------------------------------------------------------------------------

max_protein = max(max_nutrients$protein_per_gram)
highest_protein= max_nutrients %>% filter(protein_per_gram == max_protein) %>% select(Food,Grams,Carbs,Protein,Fat,carbs_per_gram,protein_per_gram,fat_per_gram)

highest_protein # food with max protein

#---------------------------------------------------------------------------------------------------

max_fat = max(max_nutrients$fat_per_gram)
highest_fat= max_nutrients %>% filter(fat_per_gram == max_fat) %>% select(Food,Grams,Carbs,Protein,Fat,carbs_per_gram,protein_per_gram,fat_per_gram)

highest_fat # food with max fat

```

library(dplyr)
category_data = unique(my_nurients_data$Category)  ## finding diffrent categories


## convering grams into calories for protein,carbs,fat

cal_data = my_nurients_data %>% mutate(Protein = 4*Protein , Carbs = 4*Carbs, Fat = 9*Fat) %>% select(Food,Calories,Grams,Protein,Carbs,Fat,Category)
cal_data

## sum of each nutrients for finding percentage of each nutrients category wise
#total_nutrients = cal_dat %>% group_by(Category) %>% summarise(sum(Grams),sum(Calories),sum(Protein),sum(Fat),sum(Carbs))

```

## recommending food
```{r}


c = cal_data %>% filter(Carbs<=micronutrients$carbs_cal) %>% select(Food,Carbs) %>% arrange(desc(Carbs))
c1 = head(c,1)


p = cal_data %>% filter(Protein<=micronutrients$protein_cal) %>% select(Food,Protein) %>% arrange(desc(Protein))
p1=head(p,1)



f = cal_data %>% filter(Fat<=micronutrients$fat_cal) %>% select(Food,Fat) %>% arrange(desc(Fat))
f1 =head(f,1)



recommended_food = data.frame("Carbs"=c1$Food,"prot"=p1$Food,"fat"=f1$Food) 
recommended_food


```
